Comparing human and model-based forecasts of COVID-19 in Germany and Poland

### 2022-08-20 manuscript/manuscript.tex
small manuscript revisions

\newcommand{\getAff}[1]{
  \getIndex{#1}{London School of Hygiene \& Tropical Medicine,Karlsruhe
Institute of Technology Institute of Economic Theory and Statistics,Max
Planck Institute for Multidisciplinary Sciences,UK Health Security
Agency,Imperial College London}
}

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Comparing human and model-based forecasts of COVID-19 in
Germany and
Poland} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Nikos I. Bosse\textsuperscript{\getAff{London School of Hygiene \&
Tropical Medicine}}\textsuperscript{*},
Sam Abbott\textsuperscript{\getAff{London School of Hygiene \& Tropical
Medicine}},
Johannes Bracher\textsuperscript{\getAff{Karlsruhe Institute of
Technology Institute of Economic Theory and Statistics}},
Habakuk Hain\textsuperscript{\getAff{Max Planck Institute for
Multidisciplinary Sciences}},
Billy J. Quilty\textsuperscript{\getAff{London School of Hygiene \&
Tropical Medicine}},
Mark Jit\textsuperscript{\getAff{London School of Hygiene \& Tropical
Medicine}},
Centre for the Mathematical Modelling of Infectious Diseases COVID-19
Working Group\textsuperscript{\getAff{London School of Hygiene \&
Tropical Medicine}},
Edwin van Leeuwen\textsuperscript{\getAff{UK Health Security
Agency}, \getAff{London School of Hygiene \& Tropical Medicine}},
Anne Cori\textsuperscript{\getAff{Imperial College London}},
Sebastian Funk\textsuperscript{\getAff{London School of Hygiene \&
Tropical Medicine}}\\
\bigskip
\textbf{\getAff{London School of Hygiene \& Tropical Medicine}}London
School of Hygiene \& Tropical Medicine, Keppel Street, London WC1E7HT,
United Kingdom\\
\textbf{\getAff{Karlsruhe Institute of Technology Institute of Economic
Theory and Statistics}}Karlsruhe Institute of Technology - Institute of
Economic Theory and Statistics, Post Box 6980, 76049 Karlsruhe,
Germany\\
\textbf{\getAff{Max Planck Institute for Multidisciplinary Sciences}}Max
Planck Institute for Multidisciplinary Sciences, Am Faßberg 11, 37077
Göttingen, Germany\\
\textbf{\getAff{UK Health Security Agency}}UK Health Security Agency,
London NW9 5EQ, United Kingdom\\
\textbf{\getAff{Imperial College London}}Imperial College London,
Norfolk Place, London W2 1PG, United Kingdom\\
\bigskip
* Corresponding author: nikos.bosse@lshtm.ac.uk\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Forecasts based on epidemiological modelling have played an important
role in shaping public policy throughout the COVID-19 pandemic. This
modelling combines knowledge about infectious disease dynamics with the
subjective opinion of the researcher who develops and refines the model
and often also adjusts model outputs. Developing a forecast model is
difficult, resource- and time-consuming. It is therefore worth asking
what modelling is able to add beyond the subjective opinion of the
researcher alone. To investigate this, we analysed different real-time
forecasts of cases of and deaths from COVID-19 in Germany and Poland
over a 1-4 week horizon submitted to the German and Polish Forecast Hub.
We compared crowd forecasts elicited from researchers and volunteers,
against a) forecasts from two semi-mechanistic models based on common
epidemiological assumptions and b) the ensemble of all other models
submitted to the Forecast Hub. We found crowd forecasts, despite being
overconfident, to outperform all other methods across all forecast
horizons when forecasting cases (weighted interval score relative to the
Hub ensemble 2 weeks ahead: 0.89). Forecasts based on computational
models performed comparably better when predicting deaths (rel. WIS
1.26), suggesting that epidemiological modelling and human judgement can
complement each other in important ways.

@@

Data on daily reported test positive cases and deaths linked to COVID-19
were provided by the organisers of the German and Polish Forecast hub.
Until December 14th, 2020, these data were sourced from the European
Centre for Disease Control {[}33{]}. After ECDC stopped publishing daily
data, observations were sourced from the Robert Koch Institute (RKI) and
the Polish Ministry of Health for the remainder of the submission period
{[}34{]}. These data are subject to reporting artefacts, (such as for
example delayed case reporting in Poland on the 24th November,
{[}35{]}), changes in reporting over time, and variation in testing
regimes (for example in Germany from the 11th of November on, {[}36{]}).
The ECDC data as well as the data published by the Polish Ministry of
Health were also subject to data revisions, although most of them (with
a notable exception of a data update for October 12 2020 in Germany)
only affected daily, not weekly data (see Figures
\ref{fig:daily-truth-update} and \ref{fig:weekly-truth-update}).

@@

Our crowd forecasts were created as an ensemble of forecasts made by
individual participants every week through a web application
(https://cmmid-lshtm.shinyapps.io/crowd-forecast/). Weekly forecasts had
to be submitted before Tuesday 12pm every week, but participants were
asked to only use any information or data that was already available by
Monday night. The application was built using the \texttt{shiny} and
\texttt{golem} R packages {[}37,38{]} and is available in the
\texttt{crowdforecastr} R package {[}31{]}. To make a forecast in the
application participants could select a predictive distribution (with
the default being log-normal) to represent the probability that the
forecasted quantity took certain values. Median and width of the
uncertainty could be adjusted by either interacting with a figure
showing their forecast or providing numerical values (see screenshot in
Figure \ref{fig:screenshot} in the SI). The default shown was a
repetition of the last known observation with constant uncertainty
around it computed as the standard deviation of the last four changes in
weekly log observed forecasts (i.e.~as
\(\sigma(log(value4) - log(value3), log(value3) - log(value2), \ldots )\)).
A comparison of the crowd forecasts against the default baseline shown
in the application is displayed in Figure \ref{fig:compare-forecasters}
in the Appendix. Our interface also allowed participants to view past
observations based on the hub data, as well as their forecasts, on a
logarithmic scale and presented additional contextual COVID-19 data
sourced from {[}39{]}. These data included, for example, notifications
of both test positive COVID-19 cases and COVID-19 linked deaths and the
number of COVID-19 tests conducted over time. From November 26 2020 on
we displayed weekly small reports with a visualisation of past forecasts
and scores on our website, epiforecasts.io.

@@

We used two Bayesian semi-mechanistic models from the \texttt{EpiNow2} R
package (version 1.3.3) as our model-based forecasts {[}40{]}. The first
of these models, here called ``renewal model'', used the renewal
equation {[}41{]} to predict reported cases and deaths (see details in
the SI). It estimated the effective reproduction number \(R_t\) (the
average number of people each person infected at time t is expected to
infect in turn) and modelled future infections as a weighted sum of past
infection multiplied by \(R_t\). \(R_t\) was assumed to stay constant
beyond the forecast date, roughly corresponding to continuing the latest
exponential trend in infections. On the 9th of November we altered the
date when \(R_t\) was assumed to be constant from two weeks prior to the
date of the forecast to the forecast date, which we found to yield a
more stable \(R_t\) estimate. Reported case and death notifications were
obtained by convolving predicted infections over data-based delay
distributions {[}40,42--44{]} to model the time between infection and
report date. The renewal model was used to predict cases as well as
deaths with forecasts being generated for each target separately. Death
forecasts from the renewal model were therefore not informed by past
cases. One submission of the renewal model on December 28th 2020 was
delayed and therefore not included in the official Forecast hub
ensemble.

The second model (``convolution model'', see details in SI) was only
used to forecast deaths and was added later, starting December 7th 2020
(with the first forecast from December 7th suffering from a software bug
and therefore disregarded in all further analyses). The convolution
model was submitted, but never included in the official Forecast hub
ensemble due to concerns that it could be too similar to the renewal
model. The convolution model predicted deaths as a fraction of infected
people who would die with some delay, by using a convolution of reported
cases with a distribution that described the delay from case report to
death and a scaling factor (the case-fatality ratio). Both the renewal
and the convolution model used daily observations and assumed a negative
binomial observation model with a multiplicative day-of-the-week effect
{[}40{]}.

@@

For the main analysis we focused mostly on two week ahead forecasts, as
COVID-19 forecasts, especially for cases, were in the past found to have
poor predictive performance beyond this horizon {[}15{]}. Forecasts for
cases were scored using the full period from October 2020 until March
2021. To ensure comparability between models, all death forecasts were
scored using only the period from December 14th on, where all models
including the convolution model were available. To ensure robustness of
our results we conducted a sensitivity analysis where all forecasts
(including cases) were scored only over the later period for which all
forecasts were available (see Section \ref{sensitivity-analysis} in the
SI). Results remained broadly unchanged.

Forecasts were analysed using the following scoring metrics: The
weighted interval score (WIS) {[}47{]}, the absolute error, relative
bias, and empirical coverage of the 50\% and 90\% prediction intervals.
The WIS is a proper scoring rule {[}48{]}, meaning that in expectation
the score is optimised by reporting a predictive distribution that is
identical to the true data-generating distribution. Forecasters are
therefore incentivised to report their true belief about the future. The
WIS can be understood as a generalisation of the absolute error to
quantile-based forecasts (also meaning that smaller values are better)
and can be decomposed into three separate penalties: forecast spread
(i.e.~uncertainty of forecasts), over-prediction and under-prediction.
While the over- and under-prediction components of the WIS capture the
amount of over-prediction and under-prediction in absolute terms, we
also look at a relative tendency to make biased forecasts. The bias
metric {[}9{]} we use captures how much probability mass of the forecast
was above or below the true value (mapped to values between -1 and 1)
and therefore represents a general tendency to over- or under-predict in
relative terms. A value of -1 implies that all quantiles of the
predictive distribution are below the observed value and a value of 1
that all quantiles are above the observed value. Empirical coverage is
the percentage of observed values that fall inside a given prediction
interval (e.g.~how many observed values fall inside all 50\% prediction
intervals). Scoring metrics are explained in more detail in Table
\ref{tab:scoring-metrics} in the SI. All scores were calculated using
the \texttt{scoringutils} R package {[}49{]}.

At all stages of the evaluation our forecasts were compared to the
median ensemble of all \emph{other} models submitted to the German and
Polish Forecast Hub (``Hub ensemble''). This ``Hub ensemble'' was
retrospectively computed and excludes all our models, leaving on average
five ensemble member models (see details in Section
\ref{contributions-hub}, as well as in Table
\ref{tab:table-ensemble-versions} and Figure
\ref{fig:num-ensemble-members} in the SI). What we call ``Hub ensemble''
in this article therefore differs from the ``official Hub ensemble''
(here called ``hub-ensemble-realised'') which included crowd forecasts
as well as renewal model forecasts. To enhance interpretability of
scores we mainly report WIS relative to the Hub ensemble in the main
text, i.e.~we divided the average scores for a given model by the
average score achieved by the Hub ensemble on the same set of forecasts
(with values \textgreater1 implying worse and values \textless1 implying
better performance than the Hub ensemble). In addition to comparing our
forecasts against the hub ensemble excluding our models, we also
assessed the impact of our forecasts on the performance of the
forecasting hub by recalculating separate versions of the Hub ensemble
with only some (or all) of our forecasts included. Versions that
included either all of our models (``hub-ensemble-with-all'') or only
one of them (``hub-ensemble-with-X'') were computed retrospectively.

@@

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/aggregate-performance-all-v4} \caption{Visualisation of aggregate performance metrics for forecasts one to four weeks into the future. A, B: mean weighted interval score (WIS, lower indicates better performance) across horizons. WIS is decomposed into its components dispersion, over-prediction and under-prediction. C: Empirical coverage of the 50\% prediction intervals (50\% coverage is perfect). D: Empirical coverage of the 90\% prediction intervals. E: Dispersion (same as in panel A, B). Higher values mean greater dispersion of the forecast and imply ceteris paribus a worse score. F: Bias, i.e. general (relative) tendency to over- or underpredict. Values are between -1 (complete under-prediction) and 1 (complete over-prediction) and 0 ideally. G: Absolute error of the median forecast (lower is better). H. Standard deviation of all WIS values for different horizons}\label{fig:agg-performance-all}
\end{figure}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/figure-forecasts-2} \caption{Two week ahead forecasts and corresponding scores. A, C: Visualisation of 50\% prediction intervals of two week ahead forecasts against the reported values. Forecasts that were not scored (because there was no complete set of death forecasts available) are greyed out. B, D: Visualisation of corresponding WIS.}\label{fig:forecasts-and-truth}
\end{figure}

@@

For cases, crowd forecasts had a lower mean weighted interval score
(WIS, lower values indicate better performance) than both the renewal
model and the Hub ensemble across all forecast horizons (Figure
\ref{fig:agg-performance-all}A) and locations (Figure
\ref{fig:performance-locations-rel}A). For two week ahead forecasts,
mean WIS relative to the Hub ensemble (= 1) was 0.89 for crowd forecasts
and 1.40 for the renewal model (Table \ref{tab:score-table-2}). Across
all forecasting approaches, locations and forecast horizons, the
distribution of WIS values was very right-skewed, and average
performance was heavily influenced by outliers (Figure
\ref{fig:distribution-scores}). Overall, low variance in forecast
performance was closely linked with good mean performance (Figures
\ref{fig:agg-performance-all}H and and \ref{fig:agg-performance-all}A),
suggesting that the ability to avoid large errors was an important
factor in determining overall performance. The impact of outlier values
was especially pronounced for the renewal model, which had more outliers
(Figure \ref{fig:distribution-scores}), as well as the highest standard
deviation of WIS values (standard deviation of the WIS relative to the
WIS sd of the Hub ensemble was 1.54 at the two weeks ahead horizon),
while the ensemble of crowd forecasts (rel. WIS sd 0.76) and the Hub
ensemble (= 1) showed more stable performance.

To varying degrees, all forecasts exhibited trend-following behaviour
and were rarely able to predict a change in trend before it had
happened. For example, all forecasts failed to predict the change in
trend from increase to decrease that happened in November in Germany and
severely overshot reported cases (Figure
\ref{fig:forecasts-and-truth}A). This was most striking for the renewal
model, which extrapolated unconstrained exponential growth based on the
recent past of observations. The Hub ensemble and the crowd forecast,
which had both been under-predicting throughout October, also failed to
predict the change in trend after cases peaked, but less severely so.
Human forecasters, possibly aware of the semi-lockdown announced on
November 2nd 2020 {[}50{]} and the change in the testing regime (with
stricter test criteria) on November 11th 2020 {[}36{]}, were fastest to
adapt to the new trend, and the Hub ensemble slowest. In December, cases
rose again in Germany, with all models under-predicting this growth to
varying extents. As in October, the renewal model captured the phase of
exponential growth in cases slightly better than other approaches, but
again overshot when reported case numbers fell over Christmas. The large
variance in predictions in January in Germany (severe under-prediction
followed by severe over-prediction) may in part be caused by the fact
that the renewal model operated on daily data and therefore was
susceptible to fluctuations in daily reporting around Christmas that
would not have influenced on weekly reporting. Similar trends in
performance were evident in Poland, with the crowd forecast quickest at
adapting to the change in trend in November. In general, there were
fewer large outlier forecasts in Poland and in particular the renewal
model performed more in line with other forecasts there.

All forecasting approaches, including the Hub ensemble, were
overconfident, i.e.~they showed lower than nominal coverage (meaning
that 50\% (90\%) prediction intervals generally covered less than 50\%
(90\%) of the actually observed values) (Figure
\ref{fig:agg-performance-all}C and \ref{fig:agg-performance-all}D).
Coverage for all forecasts deteriorated with increasing forecast
horizon, indicating that all forecasting approaches struggled to
quantify uncertainty appropriately for case forecasts. This was
especially an issue for crowd forecasts, which had markedly shorter
prediction intervals (i.e., narrower and more confident predictive
distributions) than other approaches (Figure
\ref{fig:agg-performance-all}E) and only showed a small increase in
uncertainty across forecast horizons. The crowd forecasts prediction
intervals were also noticeably narrower than the default baseline shown
to forecasters in the application (see Figure
\ref{fig:compare-forecasters}).

In spite of good performance in terms of the absolute error (Figure
\ref{fig:agg-performance-all}G), the narrow forecast intervals led to
forecasts which were severely overconfident (covering only 36\% and 55\%
of all observations with the 50\% and 90\% prediction intervals of all
forecasts made at a two week forecast horizon, and only 5\% and 38\%
four weeks ahead) (Figure \ref{fig:agg-performance-all}C,D and Tables
\ref{tab:score-table-2} and \ref{tab:score-table-4}). Despite worse
performance in terms of absolute error (Figure
\ref{fig:agg-performance-all}G), the renewal model achieved better
calibration (comparable to the Hub ensemble), as uncertainty increased
rapidly across forecast horizons. The crowd forecasts, on the other
hand, showed a smaller bias than the renewal model, but were
overconfident.

The renewal model exhibited a noticeable tendency towards
over-predicting reported cases across all horizons. The crowd forecast
tended to over-predict at longer forecast horizons, whereas the Hub
ensemble showed no systematic bias (Figure
\ref{fig:agg-performance-all}F). Regardless of a general relative
tendency to over-predict, all forecasting approaches incurred larger
absolute penalties from over- than from under-prediction (see
decomposition of the WIS into absolute penalties for over-prediction,
under-prediction and dispersion in Figures
\ref{fig:agg-performance-all}A and \ref{fig:agg-performance-all}B and
Tables \ref{tab:score-table-2} and \ref{tab:score-table-4}).

>>>>

\newcommand{\getAff}[1]{
  \getIndex{#1}{London School of Hygiene \& Tropical
Medicine,CMMID,Karlsruhe Institute of Technology Institute of Economic
Theory and Statistics,Max Planck Institute for Multidisciplinary
Sciences,UK Health Security Agency,Imperial College London}
}

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Comparing human and model-based forecasts of COVID-19 in
Germany and
Poland} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Nikos I. Bosse\textsuperscript{\getAff{London School of Hygiene \&
Tropical Medicine}, \getAff{CMMID}}\textsuperscript{*},
Sam Abbott\textsuperscript{\getAff{London School of Hygiene \& Tropical
Medicine}, \getAff{CMMID}},
Johannes Bracher\textsuperscript{\getAff{Karlsruhe Institute of
Technology Institute of Economic Theory and Statistics}},
Habakuk Hain\textsuperscript{\getAff{Max Planck Institute for
Multidisciplinary Sciences}},
Billy J. Quilty\textsuperscript{\getAff{London School of Hygiene \&
Tropical Medicine}, \getAff{CMMID}},
Mark Jit\textsuperscript{\getAff{London School of Hygiene \& Tropical
Medicine}, \getAff{CMMID}},
Centre for the Mathematical Modelling of Infectious Diseases COVID-19
Working Group\textsuperscript{\getAff{London School of Hygiene \&
Tropical Medicine}, \getAff{CMMID}},
Edwin van Leeuwen\textsuperscript{\getAff{UK Health Security
Agency}, \getAff{London School of Hygiene \& Tropical Medicine}},
Anne Cori\textsuperscript{\getAff{Imperial College London}},
Sebastian Funk\textsuperscript{\getAff{London School of Hygiene \&
Tropical Medicine}, \getAff{CMMID}}\\
\bigskip
\textbf{\getAff{London School of Hygiene \& Tropical
Medicine}}Department of Infectious Disease Epidemiology, London School
of Hygiene \& Tropical Medicine, London, United Kingdom\\
\textbf{\getAff{CMMID}}Centre for the Mathematical Modelling of
Infectious Diseases (members of the CMMID COVID-19 working group are
listed in S1 Acknowledgements), London, United Kingdom\\
\textbf{\getAff{Karlsruhe Institute of Technology Institute of Economic
Theory and Statistics}}Institute of Economic Theory and Statistics,
Karlsruhe Institute of Technology, Karlsruhe, Germany\\
\textbf{\getAff{Max Planck Institute for Multidisciplinary Sciences}}Max
Planck Institute for Multidisciplinary Sciences, Göttingen, Germany\\
\textbf{\getAff{UK Health Security Agency}}UK Health Security Agency,
London United Kingdom\\
\textbf{\getAff{Imperial College London}}MRC Centre for Outbreak
Analysis and Modelling, Department of Infectious Disease Epidemiology,
School of Public Health, Imperial College London, London, United
Kingdom\\
\bigskip
* nikos.bosse@lshtm.ac.uk\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Forecasts based on epidemiological modelling have played an important
role in shaping public policy throughout the COVID-19 pandemic. This
modelling combines knowledge about infectious disease dynamics with the
subjective opinion of the researcher who develops and refines the model
and often also adjusts model outputs. Developing a forecast model is
difficult, resource- and time-consuming. It is therefore worth asking
what modelling is able to add beyond the subjective opinion of the
researcher alone. To investigate this, we analysed different real-time
forecasts of cases of and deaths from COVID-19 in Germany and Poland
over a 1-4 week horizon submitted to the German and Polish Forecast Hub.
We compared crowd forecasts elicited from researchers and volunteers,
against a) forecasts from two semi-mechanistic models based on common
epidemiological assumptions and b) the ensemble of all other models
submitted to the Forecast Hub. We found crowd forecasts, despite being
overconfident, to outperform all other methods across all forecast
horizons when forecasting cases (weighted interval score relative to the
Hub ensemble 2 weeks ahead: 0.89). Forecasts based on computational
models performed comparably better when predicting deaths (rel. WIS
1.26), suggesting that epidemiological modelling and human judgement can
complement each other in important ways.

@@

Data on daily reported test positive cases and deaths linked to COVID-19
were provided by the organisers of the German and Polish Forecast hub.
Until December 14th, 2020, these data were sourced from the European
Centre for Disease Control {[}33{]}. After ECDC stopped publishing daily
data, observations were sourced from the Robert Koch Institute (RKI) and
the Polish Ministry of Health for the remainder of the submission period
{[}34{]}. These data are subject to reporting artefacts, (such as for
example delayed case reporting in Poland on the 24th November,
{[}35{]}), changes in reporting over time, and variation in testing
regimes (for example in Germany from the 11th of November on, {[}36{]}).
The ECDC data as well as the data published by the Polish Ministry of
Health were also subject to data revisions, although most of them (with
a notable exception of a data update for October 12 2020 in Germany)
only affected daily, not weekly data (see
\nameref{fig:daily-truth-update} and \nameref{fig:weekly-truth-update}
Figs).

@@

Our crowd forecasts were created as an ensemble of forecasts made by
individual participants every week through a web application
(https://cmmid-lshtm.shinyapps.io/crowd-forecast/). Weekly forecasts had
to be submitted before Tuesday 12pm every week, but participants were
asked to only use any information or data that was already available by
Monday night. The application was built using the \texttt{shiny} and
\texttt{golem} R packages {[}37,38{]} and is available in the
\texttt{crowdforecastr} R package {[}31{]}. To make a forecast in the
application participants could select a predictive distribution (with
the default being log-normal) to represent the probability that the
forecasted quantity took certain values. Median and width of the
uncertainty could be adjusted by either interacting with a figure
showing their forecast or providing numerical values (see screenshot in
\nameref{fig:screenshot}). The default shown was a repetition of the
last known observation with constant uncertainty around it computed as
the standard deviation of the last four changes in weekly log observed
forecasts (i.e.~as
\(\sigma(log(value4) - log(value3), log(value3) - log(value2), \ldots )\)).
A comparison of the crowd forecasts against the default baseline shown
in the application is displayed in \nameref{fig:compare-forecasters}.
Our interface also allowed participants to view past observations based
on the hub data, as well as their forecasts, on a logarithmic scale and
presented additional contextual COVID-19 data sourced from {[}39{]}.
These data included, for example, notifications of both test positive
COVID-19 cases and COVID-19 linked deaths and the number of COVID-19
tests conducted over time. From November 26 2020 on we displayed weekly
small reports with a visualisation of past forecasts and scores on our
website, epiforecasts.io.

@@

We used two Bayesian semi-mechanistic models from the \texttt{EpiNow2} R
package (version 1.3.3) as our model-based forecasts {[}40{]}. The first
of these models, here called ``renewal model'', used the renewal
equation {[}41{]} to predict reported cases and deaths (see details in
\nameref{txt:details-models}). It estimated the effective reproduction
number \(R_t\) (the average number of people each person infected at
time t is expected to infect in turn) and modelled future infections as
a weighted sum of past infection multiplied by \(R_t\). \(R_t\) was
assumed to stay constant beyond the forecast date, roughly corresponding
to continuing the latest exponential trend in infections. On the 9th of
November we altered the date when \(R_t\) was assumed to be constant
from two weeks prior to the date of the forecast to the forecast date,
which we found to yield a more stable \(R_t\) estimate. Reported case
and death notifications were obtained by convolving predicted infections
over data-based delay distributions {[}40,42--44{]} to model the time
between infection and report date. The renewal model was used to predict
cases as well as deaths with forecasts being generated for each target
separately. Death forecasts from the renewal model were therefore not
informed by past cases. One submission of the renewal model on December
28th 2020 was delayed and therefore not included in the official
Forecast hub ensemble.

The second model (``convolution model'', see details in
\nameref{txt:details-models}) was only used to forecast deaths and was
added later, starting December 7th 2020 (with the first forecast from
December 7th suffering from a software bug and therefore disregarded in
all further analyses). The convolution model was submitted, but never
included in the official Forecast hub ensemble due to concerns that it
could be too similar to the renewal model. The convolution model
predicted deaths as a fraction of infected people who would die with
some delay, by using a convolution of reported cases with a distribution
that described the delay from case report to death and a scaling factor
(the case-fatality ratio). Both the renewal and the convolution model
used daily observations and assumed a negative binomial observation
model with a multiplicative day-of-the-week effect {[}40{]}.

@@

For the main analysis we focused mostly on two week ahead forecasts, as
COVID-19 forecasts, especially for cases, were in the past found to have
poor predictive performance beyond this horizon {[}15{]}. Forecasts for
cases were scored using the full period from October 2020 until March
2021. To ensure comparability between models, all death forecasts were
scored using only the period from December 14th on, where all models
including the convolution model were available. To ensure robustness of
our results we conducted a sensitivity analysis where all forecasts
(including cases) were scored only over the later period for which all
forecasts were available (see \nameref{fig:agg-performance-all-late} and
\nameref{tab:score-table-late-2} and \nameref{tab:score-table-late-4}s).
Results remained broadly unchanged.

Forecasts were analysed using the following scoring metrics: The
weighted interval score (WIS) {[}47{]}, the absolute error, relative
bias, and empirical coverage of the 50\% and 90\% prediction intervals.
The WIS is a proper scoring rule {[}48{]}, meaning that in expectation
the score is optimised by reporting a predictive distribution that is
identical to the true data-generating distribution. Forecasters are
therefore incentivised to report their true belief about the future. The
WIS can be understood as a generalisation of the absolute error to
quantile-based forecasts (also meaning that smaller values are better)
and can be decomposed into three separate penalties: forecast spread
(i.e.~uncertainty of forecasts), over-prediction and under-prediction.
While the over- and under-prediction components of the WIS capture the
amount of over-prediction and under-prediction in absolute terms, we
also look at a relative tendency to make biased forecasts. The bias
metric {[}9{]} we use captures how much probability mass of the forecast
was above or below the true value (mapped to values between -1 and 1)
and therefore represents a general tendency to over- or under-predict in
relative terms. A value of -1 implies that all quantiles of the
predictive distribution are below the observed value and a value of 1
that all quantiles are above the observed value. Empirical coverage is
the percentage of observed values that fall inside a given prediction
interval (e.g.~how many observed values fall inside all 50\% prediction
intervals). Scoring metrics are explained in more detail in
\nameref{tab:scoring-metrics}. All scores were calculated using the
\texttt{scoringutils} R package {[}49{]}.

At all stages of the evaluation our forecasts were compared to the
median ensemble of all \emph{other} models submitted to the German and
Polish Forecast Hub (``Hub ensemble''). This ``Hub ensemble'' was
retrospectively computed and excludes all our models, leaving on average
five ensemble member models (see \nameref{tab:table-ensemble-versions}
and \nameref{fig:num-ensemble-members}). What we call ``Hub ensemble''
in this article therefore differs from the ``official Hub ensemble''
(here called ``hub-ensemble-realised'') which included crowd forecasts
as well as renewal model forecasts. To enhance interpretability of
scores we mainly report WIS relative to the Hub ensemble in the main
text, i.e.~we divided the average scores for a given model by the
average score achieved by the Hub ensemble on the same set of forecasts
(with values \textgreater1 implying worse and values \textless1 implying
better performance than the Hub ensemble). In addition to comparing our
forecasts against the hub ensemble excluding our models, we also
assessed the impact of our forecasts on the performance of the
forecasting hub by recalculating separate versions of the Hub ensemble
with only some (or all) of our forecasts included. Versions that
included either all of our models (``hub-ensemble-with-all'') or only
one of them (``hub-ensemble-with-X'') were computed retrospectively.

@@

\begin{figure}[H]
\caption{\bf{Visualisation of aggregate performance metrics for forecasts one to four weeks into the future.}}
A, B: mean weighted interval score (WIS, lower indicates better performance) across horizons. WIS is decomposed into its components dispersion, over-prediction and under-prediction. C: Empirical coverage of the 50\% prediction intervals (50\% coverage is perfect). D: Empirical coverage of the 90\% prediction intervals. E: Dispersion (same as in panel A, B). Higher values mean greater dispersion of the forecast and imply ceteris paribus a worse score. F: Bias, i.e. general (relative) tendency to over- or underpredict. Values are between -1 (complete under-prediction) and 1 (complete over-prediction) and 0 ideally. G: Absolute error of the median forecast (lower is better). H. Standard deviation of all WIS values for different horizons
\label{fig:agg-performance-all}
\end{figure}

\begin{figure}[H]
\caption{\bf{Two week ahead forecasts and corresponding scores. }}
A, C: Visualisation of 50\% prediction intervals of two week ahead forecasts against the reported values. Forecasts that were not scored (because there was no complete set of death forecasts available) are greyed out. B, D: Visualisation of corresponding WIS.
\label{fig:forecasts-and-truth}
\end{figure}

@@

For cases, crowd forecasts had a lower mean weighted interval score
(WIS, lower values indicate better performance) than both the renewal
model and the Hub ensemble across all forecast horizons (Fig
\ref{fig:agg-performance-all}A) and locations
(\nameref{fig:performance-locations-rel}A Fig). For two week ahead
forecasts, mean WIS relative to the Hub ensemble (= 1) was 0.89 for
crowd forecasts and 1.40 for the renewal model
(\nameref{tab:score-table-2} Table). Across all forecasting approaches,
locations and forecast horizons, the distribution of WIS values was very
right-skewed, and average performance was heavily influenced by outliers
(Fig \ref{fig:distribution-scores}). Overall, low variance in forecast
performance was closely linked with good mean performance (Fig
\ref{fig:agg-performance-all}H and \ref{fig:agg-performance-all}A),
suggesting that the ability to avoid large errors was an important
factor in determining overall performance. The impact of outlier values
was especially pronounced for the renewal model, which had more outliers
(Fig \ref{fig:distribution-scores}), as well as the highest standard
deviation of WIS values (standard deviation of the WIS relative to the
WIS sd of the Hub ensemble was 1.54 at the two weeks ahead horizon),
while the ensemble of crowd forecasts (rel. WIS sd 0.76) and the Hub
ensemble (= 1) showed more stable performance.

To varying degrees, all forecasts exhibited trend-following behaviour
and were rarely able to predict a change in trend before it had
happened. For example, all forecasts failed to predict the change in
trend from increase to decrease that happened in November in Germany and
severely overshot reported cases (Fig \ref{fig:forecasts-and-truth}A).
This was most striking for the renewal model, which extrapolated
unconstrained exponential growth based on the recent past of
observations. The Hub ensemble and the crowd forecast, which had both
been under-predicting throughout October, also failed to predict the
change in trend after cases peaked, but less severely so. Human
forecasters, possibly aware of the semi-lockdown announced on November
2nd 2020 {[}50{]} and the change in the testing regime (with stricter
test criteria) on November 11th 2020 {[}36{]}, were fastest to adapt to
the new trend, and the Hub ensemble slowest. In December, cases rose
again in Germany, with all models under-predicting this growth to
varying extents. As in October, the renewal model captured the phase of
exponential growth in cases slightly better than other approaches, but
again overshot when reported case numbers fell over Christmas. The large
variance in predictions in January in Germany (severe under-prediction
followed by severe over-prediction) may in part be caused by the fact
that the renewal model operated on daily data and therefore was
susceptible to fluctuations in daily reporting around Christmas that
would not have influenced on weekly reporting. Similar trends in
performance were evident in Poland, with the crowd forecast quickest at
adapting to the change in trend in November. In general, there were
fewer large outlier forecasts in Poland and in particular the renewal
model performed more in line with other forecasts there.

All forecasting approaches, including the Hub ensemble, were
overconfident, i.e.~they showed lower than nominal coverage (meaning
that 50\% (90\%) prediction intervals generally covered less than 50\%
(90\%) of the actually observed values) (Fig
\ref{fig:agg-performance-all}C and \ref{fig:agg-performance-all}D).
Coverage for all forecasts deteriorated with increasing forecast
horizon, indicating that all forecasting approaches struggled to
quantify uncertainty appropriately for case forecasts. This was
especially an issue for crowd forecasts, which had markedly shorter
prediction intervals (i.e., narrower and more confident predictive
distributions) than other approaches (Fig
\ref{fig:agg-performance-all}E) and only showed a small increase in
uncertainty across forecast horizons. The crowd forecasts prediction
intervals were also noticeably narrower than the default baseline shown
to forecasters in the application (see
\nameref{fig:compare-forecasters}).

In spite of good performance in terms of the absolute error (Fig
\ref{fig:agg-performance-all}G), the narrow forecast intervals led to
forecasts which were severely overconfident (covering only 36\% and 55\%
of all observations with the 50\% and 90\% prediction intervals of all
forecasts made at a two week forecast horizon, and only 5\% and 38\%
four weeks ahead) (Fig \ref{fig:agg-performance-all}C and
\ref{fig:agg-performance-all}D as well as \nameref{tab:score-table-2}
and \nameref{tab:score-table-4}s). Despite worse performance in terms of
absolute error (Fig \ref{fig:agg-performance-all}G), the renewal model
achieved better calibration (comparable to the Hub ensemble), as
uncertainty increased rapidly across forecast horizons. The crowd
forecasts, on the other hand, showed a smaller bias than the renewal
model, but were overconfident.

The renewal model exhibited a noticeable tendency towards
over-predicting reported cases across all horizons. The crowd forecast
tended to over-predict at longer forecast horizons, whereas the Hub
ensemble showed no systematic bias (Fig \ref{fig:agg-performance-all}F).
Regardless of a general relative tendency to over-predict, all
forecasting approaches incurred larger absolute penalties from over-
than from under-prediction (see decomposition of the WIS into absolute
penalties for over-prediction, under-prediction and dispersion in Fig
\ref{fig:agg-performance-all}A and \ref{fig:agg-performance-all}B, as
well as \nameref{tab:score-table-2} and \nameref{tab:score-table-4}s).

Generally, trends in overall performance were broadly similar across
locations (\nameref{fig:performance-locations} and
\nameref{fig:performance-locations-rel} Figs). Due to the differing
population sizes and numbers of notifications in Germany and Poland
absolute scores were difficult to compare directly. However, relative to
the Hub ensemble, the crowd forecasts performed noticeably better in
Germany than in Poland and the renewal model better in Poland than in
Germany (\nameref{fig:performance-locations-rel}A,
\nameref{fig:performance-locations-rel}G,
\nameref{fig:agg-performance-all-Germany} and
\nameref{fig:agg-performance-all-Poland} Figs).
